---
title: "b3data: Data resources for the b3verse"
sidebar:
  label: b3data
  order: 7
knit: (function(inputFile, ...) {
  knitr::knit(
    input = inputFile,
    output = gsub("Rmd$", "md", inputFile)) })
---

<!-- b3data.md is generated from b3data.Rmd Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = file.path("../../../..", "public", "guides", "b3data", "/"),
  fig.height = 4,
  dpi = 300
)
```

This guide provides an overview of the **b3data** Frictionless data package and supporting scripts, which serve as foundational resources for calculating biodiversity indicators from occurrence cubes within the **b3verse**.

Suggestion citation:

**Citing this guide:**

> Langeraert W (2025). b3data: Data resources for the b3verse. <https://docs.b-cubed.eu/guides/b3data/>

**Citing the data package:**

> Langeraert W (2025). b3data: Data resources for the b3verse [Data set]. <https://doi.org/10.5281/zenodo.15181097>

**Citing the data scripts:**

> Langeraert W (2025). Scripts used to create the b3data frictionless data package [Computer software]. <https://github.com/b-cubed-eu/b3data-scripts/>

## What is b3data?

The **b3data** data package is a curated and versioned collection of datasets, designed for use in the **b3verse** indicator calculation workflow.
It is published in accordance with the [Frictionless Data](https://frictionlessdata.io/) specifications, that allow publishing datasets in a [FAIR](https://www.go-fair.org/fair-principles/) and open manner.

To learn more about the **b3verse**, visit the [documentation site](https://docs.b-cubed.eu/guides/b3verse/), or explore related packages via the [b3verse R-universe](https://b-cubed-eu.r-universe.dev/).

The data package is published on [Zenodo](https://doi.org/10.5281/zenodo.15181097) and compiled using R code:

- **Published at**: [![DOI](https://zenodo.org/badge/DOI/10.5281/zenodo.15181097.svg)](https://doi.org/10.5281/zenodo.15181097)
- **Compiled by**: [b3data-scripts](https://github.com/b-cubed-eu/b3data-scripts)
- **Used in**: [b3verse](https://docs.b-cubed.eu/guides/b3verse/)
- **Importable in R via**:
  [frictionless](https://docs.ropensci.org/frictionless/) R package (see further)

## Getting the data

The data resources can be downloaded from the [Zenodo](https://doi.org/10.5281/zenodo.15181097) repository, but they can also be accessed directly in R using the [frictionless](https://docs.ropensci.org/frictionless/) R package:

### Step 1 — Load the frictionless R package

```{r}
# install.packages("frictionless")
library(frictionless)
```

### Step 2 — Read the package descriptor from Zenodo

The content of the data package can be consulted using `read_package()`.

```{r}
b3data_package <- read_package("https://zenodo.org/records/15211029/files/datapackage.json")
b3data_package
```

This object contains metadata and references to all resources included in the data package.

### Step 3 — Import a resource (dataset)

Tabular datasets (such as occurrence cubes) can be loaded using `read_resource()`.

```{r}
bird_cube_belgium <- read_resource(b3data_package, "bird_cube_belgium_mgrs10")
head(bird_cube_belgium)
```

For non-tabular resources like spatial or raster files, use specialized packages such as `sf` or `terra` directly.

```{r, warning=FALSE, message=FALSE}
# Load package sf
library(sf)

# Read data from repository
mgrs10_belgium <- st_read(
  "https://zenodo.org/records/15211029/files/mgrs10_refgrid_belgium.gpkg",
  quiet = TRUE
)
head(mgrs10_belgium)
```

### Step 4 — Post-processing

Once loaded, resources can be filtered, joined, or passed into **b3verse** indicator functions.

```{r, warning=FALSE, message=FALSE}
# Install b3verse packages
# pkgs <- available.packages(repos = "https://b-cubed-eu.r-universe.dev")[, "Package"]
# install.packages(pkgs, repos = "https://b-cubed-eu.r-universe.dev")

# Load packages
library(b3gbi)   # General biodiversity indicators for occurrence cubes

library(dplyr)   # Data wrangling
library(ggplot2) # Data visualisation
```

We filter out any data with minimal coordinate uncertainty larger than 10 km.

```{r}
bird_cube_belgium_clean <- bird_cube_belgium %>%
  dplyr::filter(mincoordinateuncertaintyinmeters <= 100000)
```

We join the loaded resources together and visualise the number of species per grid cell.

```{r joined-map}
bird_cube_belgium_clean %>%
  # Count number of species
  summarise(
    n_species = n_distinct(species),
    .by = mgrscode
  ) %>%
  # Add MGRS grid
  left_join(mgrs10_belgium, by = join_by(mgrscode)) %>%
  st_sf(sf_column_name = "geom", crs = st_crs(mgrs10_belgium)) %>%
  # Visualise result
  ggplot() +
    geom_sf(aes(fill = n_species)) +
    theme_minimal()
```

We now process the cleaned data cube using the **b3gbi** package (v0.4.0), which prepares the data for indicator calculations.

```{r}
bird_cube_processed <- process_cube(
  bird_cube_belgium_clean,
  cols_occurrences = "n"
)
bird_cube_processed
```

With the processed cube, we can reproduce a species richness map similar to the one created earlier.

```{r}
bird_cube_richness_map <- obs_richness_map(bird_cube_processed) 
plot(bird_cube_richness_map)
```

We can also calculate other biodiversity indicators. In the example below, we generate a time series of observed species richness. Confidence intervals are omitted here to reduce computation time.

```{r}
bird_cube_richness_ts <- obs_richness_ts(bird_cube_processed, ci_type = "none") 
plot(bird_cube_richness_ts)
```

> **Note:** The spatial and temporal patterns shown in these outputs primarily reflect the data coverage in GBIF, not actual biodiversity patterns. This tutorial is meant to demonstrate the use of the **b3data** resources and indicator workflow, and is not intended as an example of rigorous ecological analysis.

## Contributing and reporting issues

**How to contribute?**
- Before contributing, check the “Contributing Guidelines” of the [b3data-scripts repository](https://github.com/b-cubed-eu/b3data-scripts/blob/main/.github/CONTRIBUTING.md).
- Fork the b3data-scripts repository.
- Follow the instructions in the README to set up the workflow.
- Submit a pull request for your changes.
- Submission of (a) resource(s) to the data package can also be requested in an issue in the repository and added by the maintainers of the data package itself.

**Reporting bugs or suggesting improvements**
- If you encounter a problem or like to suggest an improvement, open an issue in the [b3data-scripts repository](https://github.com/b-cubed-eu/b3data-scripts/issues).
- Be as detailed as possible when describing the issue, including R session info, error messages, and reproducible examples if applicable.
